---
title: "Data Description/Exploration"
author: "Eleanor and Liam"
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
editor: source
---
```{r}
#| label: setup
#| include: false

library(MASS)
library(StreamMetabolism) ## Has a very useful sunrise/sunset function.
library(tidyverse)
library(lunar)
library(GGally)
library(patchwork)
```

# Helper Functions

The `nightdater` function created rounds early-morning datetimes back to the previous night, creating a *night of* date. For example, a bird tagged at 23:00 on July 15 and a bird tagged at 3:00 on July 16 would both be considered tagged on the *night of* July 15

```{r}
#| label: helper_functions

nightdater <- function (colIn) {
  nightdate <- as.Date(trunc(colIn - ifelse(lubridate::hour(colIn) < 8, 86400, 0), by = "days"))
  return(nightdate)
}
```


# Weather Data

Our weather data from 2025 was collected via Kestrel 5500 weather meters. We had one on Gull Island, one on Great Island, and one on Baccalieu Island. We had persistent problems with the wind-measuring propellers breaking or getting jammed, so our record for wind data in particular is incomplete. One goal is to compare the island weather data with publicly available weather data from St. John's International Airport to assess whether airport data can replace missing island-level measurements.

| Island | Latitude | Longitude | Kestrel Marked Coordinates | Notes |
|:---:|:---:|:---:|:---:|:---:|
| Gull Island | 47.2600 | -52.7757 | N47 15.6, W52 46.54 | Kestrel location |
| Great Island | 47.1877 | -52.8172 | N47 11.26, W52 49.03 | Same location as light meter |
| Baccalieu Island | 48.1220 | -52.7982 | N48 7.32, W52 47.89 | Same location as light meter |

: Kestrel Weather Station Locations {tbl-colwidths="[15, 15, 15, 35, 20]"}

## Standardization of Units for Comparison

In order for proper comparison to occur some values will have to be converted to different units due to differences in standards established by the lab for the Kestrel 5500 and the St. John's International Airport. From here on out, the standard is as follows.

|Variable|Unit|
|-----|-----|
|`datetime`|YYYY-MM-DD hh:mm:ss|
|`temperature`|°C|
|`station_pressure`|kPa|
|`barometric_pressure`|kPa|
|`wind_speed`|km/h|
|`wind_chill`|°C|
|`relative_humidity`|%|
|`dew_point`|°C|

:Standardized Units of Measurements {tbl-colwidths="[75,25]"}

## Kestrel Import Function

Read in the data so the date and time readings are datetime objects and in Newfoundland time. 

The Gull Island and Great Island kestrels had long periods where their wind impellers were broken. We also typically turned the kestrels on before they were deployed and did not turn them off until several days after retrieval, so I will trim the data to remove the extra days, and replace the '0' wind speed with NA where the impellers were broken. 

```{r}
#| label: kestrel_weather_import_function

read_kestrel_weather = function(file, island, start_date, end_date, wind_all_missing = FALSE, wind_missing_after = NULL) {
  kestrel_weather = read_csv(file, show_col_types = FALSE) %>%
    mutate(
      datetime = as_datetime(Time, format = "%m/%d/%Y %H:%M"),
      datetime = force_tz(datetime, tzone = "America/St_Johns")
    ) %>%
    filter(datetime >= ymd(start_date, tz = "America/St_Johns"),
           datetime <= ymd(end_date, tz = "America/St_Johns"))
  
  if(wind_all_missing) {
    kestrel_weather = kestrel_weather %>%
      mutate(`Wind Speed` = NA_real_)
  }
  if(!is.null(wind_missing_after)) {
    kestrel_weather = kestrel_weather %>%
      mutate(`Wind Speed` = if_else(datetime > ymd(wind_missing_after), NA_real_, `Wind Speed`))
  }
  
  kestrel_weather %>%
    transmute(
      datetime,
      temperature = Temp,
      station_pressure = `Station P.` / 10,
      barometric_pressure = Baro. / 10,
      wind_speed = `Wind Speed` * 3.6,
      wind_chill = `Wind Chill`,
      relative_humidity = `Rel. Hum.`,
      dew_point = `Dew Point`,
      weather_station_location = island
    )
}
```

## Read in Kestrel data

```{r}
#| label: kestrel_data

baccalieu_weather = read_kestrel_weather(
  file = "data/baccalieu_2025_weather.csv",
  island = "Baccalieu Island",
  start_date = "2025-06-10",
  end_date = "2025-10-21"
)

gull_weather = read_kestrel_weather(
  file = "data/gull_2025_weather.csv",
  island = "Gull Island",
  start_date = "2025-06-16",
  end_date = "2025-10-29",
  wind_all_missing = TRUE
)

great_weather = read_kestrel_weather(
  file = "data/great_2025_weather.csv",
  island = "Great Island",
  start_date = "2025-06-23",
  end_date = "2025-09-05",
  wind_missing_after = "2025-07-19"
)
```
# Light Data

Our light meters are Unihedron SQM-LU-DL (Sky Quality Meter Lens USB DataLogger) light meters, made for measuring nighttime light pollution. They measure light in Magnitudes per Arcsecond Squared, which is an astronomical measurement; magnitudes relates to star brightness (the sun's magnitude is -27, Sirius has a magnitude of -1, Polaris has a magnitude of 2). Arc-seconds relate to how the night sky is divided into minutes and seconds; a square arc-second is a one-second by one-second area of the sky. If you could divide the light of a star with a magnitude of 20 (very dim; Pluto's largest moon has a magnitude of 16) evenly over one square arcsecond, the MPSAS reading would be 20. The scale is negative, with higher values being darker. A sky with an MSAS reading of 5 (the cutoff for our meters) would be around sunset. Our night skies on dark nights are typically around 21 MPSAS, and the darkest night skies on earth read around 25. Each increase of 1 MPSAS means 2.5 times less light is coming from the sky; a sky measured at 5 MPSAS is 100 times brighter than a sky measured at 10 MPSAS. 

We had one light meter on Great Island (for Witless Bay) in 2025 and one on Baccalieu. Not much data trimming needed here; I've done some pre-processing of the data already, and the light meters have been much less finicky than the weather meters. 

```{r}
#| label: light_data

baccalieu_light <- read_csv("data/baccalieu_light_20251023.csv", show_col_types = FALSE) %>%
  mutate(site = "Baccalieu Island")

witless_light <- read_csv("data/witless_light_20250913.csv", show_col_types = FALSE) %>%
  mutate(site = "Great Island, Witless Bay")

light_data = bind_rows(baccalieu_light, witless_light)
```

One thing that may prove useful is filtering the light data for before/after (etc) nautical or astronomical twilight. The data frames have a sun elevation variable; nautical night begins when the sun is 12 degrees below the horizon, and astronomical night begins when the sun is 18 degrees below the horizon.

## Sunrise and Sunset times

For doing anything with light data, it can be useful to have a list of sunrise and sunset times. Our islands are not so far apart that these need to be calculated separately. The package "StreamMetabolism" has a very useful function for calculating sunrise and sunset as new columns using a date column. However, the function is slow and computationally heavy, so I usually make a list of sunrise and sunset times for our study period, then join that list to other dataframes as needed.

```{r}
#| label: sunrise_and_sunset_times

suntimes  = tibble(
  date = seq(ymd("2025-06-01"), ymd("2025-10-31"), by = "1 day"),
  lat = 47.58,
  lon = -52.71
) %>% 
  rowwise() %>% 
  mutate(sunrise.set(lat = lat, long = lon, date = date + lubridate::days(1), timezone = "America/St_Johns")) ## Adding one day is necessary to get sunset for the given date rather than the night before
```

# St. John's International Airport Weather Data Import and Wrangling

No Barometric Pressure is recorded, only station pressure. i will include the column only for comparisons sake. 
the elevation is taken from the same website as the data was taken from. 

because the airport data is taken every hour on the half hour, the time will be rounded to match the data collected from the kestrel 5500 stations.

```{r}
#| label: SJIA_weather_data_&_matched_weather_variable_data_set
file_list = list.files("SJIA_weather_data")

airport_weather_data = as_tibble(read.csv(paste0("SJIA_weather_data/", file_list[1])))

for(i in 2:length(file_list)) {
  temp_weather_data = as_tibble(read.csv(paste0("SJIA_weather_data/", file_list[i])))
  
airport_weather_data = bind_rows(airport_weather_data, temp_weather_data)
}

master_weather_data = airport_weather_data %>%
  mutate(datetime = as_datetime(Date.Time..LST., format = "%Y-%m-%d %H:%M"),
         datetime = force_tz(datetime, tzone = "America/St_Johns"),
         datetime = round_date(datetime, "hour")) %>%
  transmute(
    datetime,
    temperature = Temp...C.,
    station_pressure = Stn.Press..kPa.,
    barometric_pressure = station_pressure * ((1 - (0.0065 * 140.50) / (temperature + 273.15 + (0.0065 * 140.50)))^-5.257),
    wind_speed = Wind.Spd..km.h.,
    wind_chill = Wind.Chill,
    relative_humidity = Rel.Hum....,
    dew_point = Dew.Point.Temp...C.,
    weather_station_location = "St. John's International Airport"
  ) %>%
  filter(datetime >= "2025-06-10" & datetime <= "2025-10-29") %>%
  bind_rows(baccalieu_weather, gull_weather, great_weather)
```

# Creation of Daily Summaries

## Daily Summary Function

```{r}
#| label: daily_summary_function

make_daily_summary = function(data, variable) {
  data %>%
    mutate(date = as_date(datetime)) %>%
    group_by(date, weather_station_location) %>%
    summarise(
      value = mean({{ variable }}, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    pivot_wider(names_from = weather_station_location,
                values_from = value)
}
```

## Creation of Daily Summary Data Frames

```{r}
#| label: daily_summary_data_frames

daily_temperature = make_daily_summary(master_weather_data, temperature)
daily_pressure = make_daily_summary(master_weather_data, barometric_pressure)
daily_wind = make_daily_summary(master_weather_data, wind_speed)
daily_humidity = make_daily_summary(master_weather_data, relative_humidity)
daily_dew_point = make_daily_summary(master_weather_data, dew_point)
```

# Correlation Matrices

## Functions for ggpairs plots

```{r}
#| label: ggpairs_functions

ggpairs_lower_function = function(data, mapping) {
  ggplot(data = data, mapping = mapping) +
    geom_point(alpha = 0.5) +
    geom_smooth(method = "lm", se = FALSE)
}

ggpairs_diagonal_function = function(data, mapping) {
  ggplot(data = data, mapping = mapping) +
    geom_density()
}
```

## Daily Temperature Correlation Matrix

```{r}
#| label: daily_temperature_correlation_matrix
#| warning: false
#| message: false

ggpairs(
  daily_temperature,
  columns = 2:ncol(daily_temperature),
  title = "Daily Mean Temperature Correlations",
  lower = list(continuous = ggpairs_lower_function),
  diag = list(continuous = ggpairs_diagonal_function)
)
```

## Daily Pressure Correlation Matrix

```{r}
#| label: daily_pressure_correlation_matrix
#| warning: false
#| message: false

ggpairs(
  daily_pressure,
  columns = 2:ncol(daily_pressure),
  title = "Daily Mean Pressure Correlations",
  lower = list(continuous = ggpairs_lower_function),
  diag = list(continuous = ggpairs_diagonal_function)
)
```

## Daily Wind Correlation Matrix

```{r}
#| label: daily_wind_correlation_matrix
#| warning: false
#| message: false

daily_wind_no_gull = daily_wind %>%
  select(-`Gull Island`)

ggpairs(
  daily_wind_no_gull,
  columns = 2:ncol(daily_wind_no_gull),
  title = "Daily Mean Wind Correlations",
  lower = list(continuous = ggpairs_lower_function),
  diag = list(continuous = ggpairs_diagonal_function)
)
```

Wind speed showed weaker agreement among stations than temperature, pressure, humidity, or dew point. Gull Island was excluded because the wind impeller failed for the full deployment. These preliminary results suggest that airport wind speed may not be a reliable substitute for island-level wind measurements, although this should be evaluated more, formally.

## Daily Humidity Correlation Matrix

```{r}
#| label: daily_humidity_correlation_matrix
#| warning: false
#| message: false

ggpairs(
  daily_humidity,
  columns = 2:ncol(daily_humidity),
  title = "Daily Mean Humidity Correlations",
  lower = list(continuous = ggpairs_lower_function),
  diag = list(continuous = ggpairs_diagonal_function)
)
```

## Daily Dew Point Correlation Matrix

```{r}
#| label: daily_dew_point_correlation_matrix
#| warning: false
#| message: false

ggpairs(
  daily_dew_point,
  columns = 2:ncol(daily_dew_point),
  title = "Daily Mean Dew Point Correlations",
  lower = list(continuous = ggpairs_lower_function),
  diag = list(continuous = ggpairs_diagonal_function)
)
```

# Weather Data for Building Models

```{r}
#| label: cleaned_airport_weather_data

cleaned_airport_weather_data = airport_weather_data %>%
  mutate(datetime = as_datetime(Date.Time..LST., format = "%Y-%m-%d %H:%M"),
         datetime = force_tz(datetime, tzone = "America/St_Johns"),
         datetime = round_date(datetime, "hour")) %>%
  select(datetime,Weather, Visibility..km., Temp...C., Stn.Press..kPa., Rel.Hum...., Hmdx, Dew.Point.Temp...C., Wind.Spd..km.h., Wind.Dir..10s.deg., Wind.Chill) %>%
  rename(
    weather = Weather,
    visibility = Visibility..km.,
    temperature = Temp...C.,
    station_pressure = Stn.Press..kPa.,
    relative_humidity = Rel.Hum....,
    humidex = Hmdx,
    dew_point = Dew.Point.Temp...C.,
    wind_speed = Wind.Spd..km.h.,
    wind_direction = Wind.Dir..10s.deg.,
    wind_chill = Wind.Chill
  ) %>%
  mutate(barometric_pressure = (station_pressure)*((1-(0.0065*140.50)/(temperature+273.15+(0.0065*140.50)))^-5.257)) %>%
  relocate(barometric_pressure, .before = relative_humidity) %>%
  mutate(weather = tolower(weather))
```

# Linear Models for Island Temperature

## Baccalieu Island Temperature As Predicted by St. John's International Airport

```{r}
#| label: baccalieu_island_airport_temperature_model

baccalieu_temperature_model = lm(
  `Baccalieu Island` ~ `St. John's International Airport`,
  data = daily_temperature
)

summary(baccalieu_temperature_model)

par(mfrow = c(2, 2))
plot(baccalieu_temperature_model)
par(mfrow = c(1, 1))
```

## Gull Island Temperature As Predicted by St. John's International Airport

```{r}
#| label: gull_island_airport_temperature_model

gull_temperature_model = lm(
  `Gull Island` ~ `St. John's International Airport`,
  data = daily_temperature
)

summary(gull_temperature_model)

par(mfrow = c(2, 2))
plot(gull_temperature_model)
par(mfrow = c(1, 1))
```

## Great Island Temperature as Predicted by St. John's International Airport

```{r}
#| label: great_island_airport_temperature_model

great_temperature_model = lm(
  `Great Island` ~ `St. John's International Airport`,
  data = daily_temperature
)

summary(great_temperature_model)

par(mfrow = c(2, 2))
plot(great_temperature_model)
par(mfrow = c(1, 1))
```

# Linear Models for Island Pressure

## Baccalieu Island Pressure As Predicted by St. John's International Airport

```{r}
#| label: baccalieu_island_airport_pressure_model

baccalieu_pressure_model = lm(
  `Baccalieu Island` ~ `St. John's International Airport`,
  data = daily_pressure
)

summary(baccalieu_pressure_model)

par(mfrow = c(2, 2))
plot(baccalieu_pressure_model)
par(mfrow = c(1, 1))
```

## Gull Island Pressure As Predicted by St. John's International Airport

```{r}
#| label: gull_island_airport_pressure_model

gull_pressure_model = lm(
  `Gull Island` ~ `St. John's International Airport`,
  data = daily_pressure
)

summary(gull_pressure_model)

par(mfrow = c(2, 2))
plot(gull_pressure_model)
par(mfrow = c(1, 1))
```

## Great Island Pressure as Predicted by St. John's International Airport

```{r}
#| label: great_island_airport_pressure_model

great_pressure_model = lm(
  `Great Island` ~ `St. John's International Airport`,
  data = daily_pressure
)

summary(great_pressure_model)

par(mfrow = c(2, 2))
plot(great_pressure_model)
par(mfrow = c(1, 1))
```

# Linear Models for Island Wind Speed

## Baccalieu Island Wind Speed As Predicted by St. John's International Airport

```{r}
#| label: baccalieu_island_airport_wind_speed_model

baccalieu_wind_model = lm(
  `Baccalieu Island` ~ `St. John's International Airport`,
  data = daily_wind
)

summary(baccalieu_wind_model)

par(mfrow = c(2, 2))
plot(baccalieu_wind_model)
par(mfrow = c(1, 1))
```

## Great Island Wind Speed as Predicted by St. John's International Airport

```{r}
#| label: great_island_airport_wind_speed_model

great_wind_model = lm(
  `Great Island` ~ `St. John's International Airport`,
  data = daily_wind
)

summary(great_wind_model)

par(mfrow = c(2, 2))
plot(great_wind_model)
par(mfrow = c(1, 1))
```

# Linear Models for Island Humidity

## Baccalieu Island Humidity As Predicted by St. John's International Airport

```{r}
#| label: baccalieu_island_airport_humidity_model

baccalieu_humidity_model = lm(
  `Baccalieu Island` ~ `St. John's International Airport`,
  data = daily_humidity
)

summary(baccalieu_humidity_model)

par(mfrow = c(2, 2))
plot(baccalieu_humidity_model)
par(mfrow = c(1, 1))
```

## Gull Island Humidity As Predicted by St. John's International Airport

```{r}
#| label: gull_island_airport_humidity_model

gull_humidity_model = lm(
  `Gull Island` ~ `St. John's International Airport`,
  data = daily_humidity
)

summary(gull_humidity_model)

par(mfrow = c(2, 2))
plot(gull_humidity_model)
par(mfrow = c(1, 1))
```

## Great Island Humidity as Predicted by St. John's International Airport

```{r}
#| label: great_island_airport_humidity_model

great_humidity_model = lm(
  `Great Island` ~ `St. John's International Airport`,
  data = daily_humidity
)

summary(great_humidity_model)

par(mfrow = c(2, 2))
plot(great_humidity_model)
par(mfrow = c(1, 1))
```

# Linear Models for Island Dew Point

## Baccalieu Island Dew Point As Predicted by St. John's International Airport

```{r}
#| label: baccalieu_island_airport_dew_point_model

baccalieu_dew_point_model = lm(
  `Baccalieu Island` ~ `St. John's International Airport`,
  data = daily_dew_point
)

summary(baccalieu_dew_point_model)

par(mfrow = c(2, 2))
plot(baccalieu_dew_point_model)
par(mfrow = c(1, 1))
```

## Gull Island Dew Point As Predicted by St. John's International Airport

```{r}
#| label: gull_island_airport_dew_point_model

gull_dew_point_model = lm(
  `Gull Island` ~ `St. John's International Airport`,
  data = daily_dew_point
)

summary(gull_dew_point_model)

par(mfrow = c(2, 2))
plot(gull_dew_point_model)
par(mfrow = c(1, 1))
```

## Great Island Dew Point as Predicted by St. John's International Airport

```{r}
#| label: great_island_airport_dew_point_model

great_dew_point_model = lm(
  `Great Island` ~ `St. John's International Airport`,
  data = daily_dew_point
)

summary(great_dew_point_model)

par(mfrow = c(2, 2))
plot(great_dew_point_model)
par(mfrow = c(1, 1))
```

# Fog Prediction Models

## Fog Presence Data

```{r}
#| label: fog_presence_data_v1

fog_data = cleaned_airport_weather_data %>%
  arrange(datetime) %>%
  filter(datetime >= "2025-06-10" & datetime <= "2025-10-28") %>%
  mutate(fog = str_detect(weather, "fog"),
         fog = replace_na(fog, FALSE),
         dew_point_depression = temperature - dew_point,
         barometric_pressure_12_hour_lag = lag(barometric_pressure, 12),
         barometric_pressure_12_hour_trend = barometric_pressure - barometric_pressure_12_hour_lag,
         date = as_date(datetime),
         fog_next_3h = lead(fog, 3)) %>%
  left_join(suntimes, by = "date") %>%
  filter(datetime >= sunset | datetime <= sunrise)
```

## Binomial Distribution Model

```{r}
#| label: binomial_distribution_model

binomial_fog_model = glm(
  fog ~ dew_point_depression +
    relative_humidity +
    visibility +
    barometric_pressure_12_hour_trend +
    wind_speed,
  family = binomial,
  data = fog_data
)

summary(binomial_fog_model)

par(mfrow = c(2, 2))
plot(binomial_fog_model)
par(mfrow = c(1, 1))
```

# Evaluation of St. John's International Airport Data For Island Approximation

## Evaluation Function

```{r}
#| label: evaluation_function

evaluate_airport_prediction = function(data, island_name) {
  model_data = data %>%
    select(
      date,
      island_value = all_of(island_name),
      airport_value = `St. John's International Airport`
    ) %>%
    drop_na()
  
  model = lm(island_value ~ airport_value, data = model_data)
  
  tibble(
    island = island_name,
    n_days = nrow(model_data),
    r = cor(model_data$island_value, model_data$airport_value),
    r_squared = summary(model)$r.squared,
    root_mean_sqaure_error = sqrt(mean(resid(model)^2)),
    mean_absolute_error = mean(abs(resid(model))),
    bias_airport_minus_island = mean(model_data$airport_value - model_data$island_value),
    intercept = coef(model)[1],
    slope = coef(model)[2]
  )
}
```

## Function to Plot Airport Against Island Data

```{r}
#| label: airport_island_plot_function

plot_airport_prediction = function(data, island_name, y_label) {
  data %>%
    select(
      date,
      island_value = all_of(island_name),
      airport_value = "St. John's International Airport",
    ) %>%
    drop_na() %>%
    ggplot(aes(x = airport_value, y = island_value)) +
    geom_point(aes(colour = date), alpha = .7) +
    geom_smooth(method = "lm", se = TRUE) +
    labs(
      x = "St. John's International Airport",
      y = y_label,
      color = "Date"
    )
}
```

## Creation of Evaluation Data Frames

### Temperature

```{r}
#| label: temperature_model_evaluation

temperature_prediction_results = bind_rows(
  evaluate_airport_prediction(daily_temperature, "Baccalieu Island"),
  evaluate_airport_prediction(daily_temperature, "Gull Island"),
  evaluate_airport_prediction(daily_temperature, "Great Island")
)
```

### Pressure

```{r}
#| label: pressure_model_evaluation

pressure_prediction_results = bind_rows(
  evaluate_airport_prediction(daily_pressure, "Baccalieu Island"),
  evaluate_airport_prediction(daily_pressure, "Gull Island"),
  evaluate_airport_prediction(daily_pressure, "Great Island")
)
```

### Wind

```{r}
#| label: wind_model_evaluation

wind_prediction_results = bind_rows(
  evaluate_airport_prediction(daily_wind, "Baccalieu Island"),
  evaluate_airport_prediction(daily_wind, "Great Island")
)
```

### Humidity

```{r}
#| label: humidity_model_evaluation

humidity_prediction_results = bind_rows(
  evaluate_airport_prediction(daily_humidity, "Baccalieu Island"),
  evaluate_airport_prediction(daily_humidity, "Gull Island"),
  evaluate_airport_prediction(daily_humidity, "Great Island")
)
```

### Dew Point

```{r}
#| label: dew_point_model_evaluation

dew_point_prediction_results = bind_rows(
  evaluate_airport_prediction(daily_dew_point, "Baccalieu Island"),
  evaluate_airport_prediction(daily_dew_point, "Gull Island"),
  evaluate_airport_prediction(daily_dew_point, "Great Island")
)
```

## Combination of Evaluation Data Frames

```{r}
#| label: master_evaluation_data_frame

airport_prediction_results = bind_rows(
  temperature_prediction_results %>%
    mutate(variable = "temperature"),
  pressure_prediction_results %>%
    mutate(variable = "barometric_pressure"),
  wind_prediction_results %>%
    mutate(variable = "wind_speed"),
  humidity_prediction_results %>%
    mutate(variable = "humidity"),
  dew_point_prediction_results %>%
    mutate(variable = "dew_point")
) %>%
  relocate(c(variable, n_days, slope, intercept), .after = island)
```

# Prediction of Fog on Islands From Light Data

## Nightly Summaries for Island Light Data

```{r}
#| label: nightly_island_light_summary

nightly_light = light_data %>%
  left_join(suntimes, by = c("date")) %>%
  select(-c(lat, lon, voltage, temp)) %>%
  filter(sunelev_deg <= -12) %>%
  group_by(site, nightdate) %>%
  summarise(
    mean_light_msas = mean(light_reading_MSAS, na.rm = TRUE),
    sd_light_msas = sd(light_reading_MSAS, na.rm = TRUE),
    median_light_msas = median(light_reading_MSAS, na.rm = TRUE),
    min_light_msas = min(light_reading_MSAS, na.rm = TRUE),
    max_light_msas = max(light_reading_MSAS, na.rm = TRUE),
    mean_moon_elevation = mean(moonelev_deg, na.rm = TRUE),
    mean_moon_illumination = mean(moonillumination, na.rm = TRUE),
    n_light_records = n(),
    .groups = "drop"
  )
```

## Fog Presence Data

As per the government of Canada's website of frequently asked questions about the historical weather data [which can be found here](https://climate.weather.gc.ca/FAQ_e.html?open=Q5) the missing data in the weather column is when there wasn't anything to report, thus the observations in which there are missing values can be treated as `FALSE` for the logical column of if fog was present.

```{r}
#| label: fog_presence_data

nightly_airport_weather = cleaned_airport_weather_data %>%
  arrange(datetime) %>%
  filter(datetime >= "2025-06-10" & datetime <= "2025-10-28") %>%
  mutate(nightdate = nightdater(datetime),
         fog = str_detect(weather, "fog"),
         fog = replace_na(fog, FALSE),
         dew_point_depression = temperature - dew_point,
         date = as_date(datetime)) %>%
  left_join(suntimes, by = "date") %>%
  filter(datetime >= sunset | datetime <= sunrise) %>%
  group_by(nightdate) %>%
  summarise(
    fog_night = any(fog),
    mean_temperature = mean(temperature, na.rm = TRUE),
    mean_relative_humidity = mean(relative_humidity, na.rm = TRUE),
    mean_visibility = mean(visibility, na.rm = TRUE),
    min_visibility = min(visibility, na.rm = TRUE),
    mean_dew_point_depression = mean(dew_point_depression, na.rm = TRUE),
    min_dew_point_depression = min(dew_point_depression, na.rm = TRUE),
    mean_wind_speed = mean(wind_speed, na.rm = TRUE),
    mean_barometric_pressure = mean(barometric_pressure, na.rm = TRUE),
    .groups = "drop"
  )
```

## Fog Light Model Data

```{r}
#| label: fog_light_model_data

fog_light_model_data = nightly_light %>%
  left_join(nightly_airport_weather, by = "nightdate")
```

## Fog Light Model

```{r}
#| label: fog_light_model

fog_light_model = glm(
  fog_night ~ #mean_relative_humidity +
    min_visibility +
    #mean_dew_point_depression +
    #mean_wind_speed +
    #mean_barometric_pressure +
    mean_light_msas +
    mean_light_msas*min_visibility +
    #mean_moon_illumination +
    site,
  family = "binomial",
  data = fog_light_model_data
)

summary(fog_light_model)

par(mfrow = c(2, 2))
plot(fog_light_model)
par(mfrow = c(1, 1))
```







